Correlation among smart grid actors

### 2015-05-22 Article.tex
update

Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences. A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Rosas-Casals2007} \cite{Pahwa}. As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance. These metrics enable scientists to get a better understanding of the structure of such complex systems. It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks. That is, they exhibit different characteritics, suggesting that they may behave differently \cite{Rosas-Casals2007}. A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities. Therefore, properties like power law degree distributions, prerential attachment, or hub oriented attacks do not necessarily apply with power grids.

Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results \cite{Pahwa}. Power grids are ideed dynamical systems governed by physical laws. Concepts such as shortest path or centrality metrics have thus to be extended to take into account the particular nature of power grids. Besides, stability and resilience of such systems are much more complex than the simple connectivity of the underlying graph \cite{Pahwa}. It is well-known for instance that frequency synchronization in the grid is a necessary condition for stability. A common solution consists in modeling the grid as a network of coupled oscillators governed by a swing equation dynamic. It can then be shown that the topology of the network impacts directly the capacity of the oscillators to reach the synchronized state.

Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently. As there exists many good articles presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point. In this paper, we rather concentrate on a different approach to the system stability. We consider a set of agents that are equiped with distributed energy ressources (DER) as well as electrical loads, such that an agent can produce and consume energy at any time. His production can be, of course, used to meet his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production to the grid. Such an agent model is known as a "\textit{prosumer}" (and will be called accordingly in this paper).

@@

The intermittent nature of renewables such as wind or solar power introduces new challenges in control design. On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present. Unfortunately, these moments do not necessarily coincide with the consumption peak hours. A first approach to remedy this situation by acting on the consumption component is the use of dynamic ectricity prices sent to the end users. Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles. In periods where the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods. On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.

>>>>

Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences. A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Pagani2012} \cite{Rosas-Casals2007} \cite{Pahwa}. As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance. These metrics enable scientists to get a better understanding of the structure of such complex systems \cite{Pagani2012}. It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks. That is, they exhibit different characteritics, suggesting that they may behave differently \cite{Rosas-Casals2007}. A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities. Therefore, properties like power law degree distributions, prerential attachment, or hub oriented attacks do not necessarily apply with power grids.

Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results \cite{Pahwa}. Power grids are ideed dynamical systems governed by physical laws. Concepts such as shortest path or centrality metrics have thus to be extended to take into account the particular nature of power grids. Besides, stability and resilience of such systems are much more complex than the simple connectivity of the underlying graph \cite{Pahwa}. It is well-known for instance that frequency synchronization in the grid is a necessary condition for stability. A common solution consists in modeling the grid as a network of coupled oscillators governed by a swing equation dynamic \cite{Filatrella2008}. It can then be shown that the topology of the network impacts directly the capacity of the oscillators to reach the synchronized state \cite{Dorfler2013}.

Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently. As there exists many good articles \cite{Ramchurn} presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point. In this paper, we rather concentrate on a different approach to the system stability. We consider a set of agents that are equiped with distributed energy ressources (DER) as well as electrical loads, such that an agent can produce and consume energy at any time. His production can be, of course, used to meet his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production to the grid. Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).

@@

The intermittent nature of renewables such as wind or solar power introduces new challenges in control design. On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present. Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}. A first approach to remedy this situation by acting on the consumption component is the use of dynamic ectricity prices sent to the end users. Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles \cite{Milligan2010}. In periods where the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods. On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.
